Discrepancy, Coresets, and Sketches in Machine Learning

### 2019-05-31 COLT_camera_ready/karninliberty19.tex
adding some references

We note that while we obtain a universal {\it additive guarantee} it is often much harder to get a  {\it multiplicative guarantee}.
For Logistic regression, for example, a recent paper \cite{DBLP:journals/corr/abs-1805-08571} provides a coreset with a  {\it multiplicative guarantee} that is based on an average sensitivity property of the dataset. 
They provide a lower bound for the size of a multiplicative error coreset proving in particular that in general, it is not possible to achieve $m \ll n$. 
The coreset they build is of cardinality $m \approx \mu\sqrt{nd^3}/\eps^2$ where $\mu \geq 1$ is a complexity measure of the dataset.  
\cite{tolochinsky2018coresets} give a generic multiplicative coreset construction for any monotonic function with $\ell_2^2$ regularization. 
The dependence they get is $\tilde O(d/\eps^2)$ ignoring logarithmic factors. 
Additive approximation coresets are also studies in the $\eps$-approximation literature which is also related to the discrepancy of the problem \cite{DBLP:journals/corr/MustafaV17}.

@@

Theorems~\ref{thm:streaming11} and~\ref{thm:streaming21} are achieved by deterministic algorithms. 
They could be thought of extensions of the MRL algorithm \cite{MRL} for streaming quantile sketching. 
Quantile sketching falls into this framework since it corresponds to $f(x,q) = 1$ if $x>q$ and $0$ else. 
More details and the proof of correctness are given in Appendix ~\ref{app:sketch proof}.

Recently, {\cite{DBLP:conf/focs/KarninLL16} provided an improved (optimal) streaming quantile coreset algorithm. 
Generalizing the construction at  \cite{DBLP:conf/focs/KarninLL16} requires more work. 
The main ideas are as follows. 
We argued above that $\tilde F_{+}$ and $\tilde F_{-}$ are both good approximations for $F$. 
We can also take $\tilde F_{\pm}$ which is $\tilde F_{+}$ or $\tilde F_{-}$ with equally probability. 
Clearly $|\tilde F_{\pm} - F| \le |E|$ as before. But now, $\E[\tilde F_{\pm}] = F$ as well. 
In the streaming algorithm, we apply this compaction (converting $F$ to $\tilde F_{\pm}$) many times to small subsets of items from the stream. This allows us to use concentration results to bound the overall error. 
The main departure from \cite{DBLP:conf/focs/KarninLL16} is that $\tilde F_{\pm}$ has half the support of $F$ only in expectation.

>>>>

We note that while we obtain a universal {\it additive guarantee} it is often much harder to get a  {\it multiplicative guarantee}.
For Logistic regression, for example, a recent paper \cite{DBLP:journals/corr/abs-1805-08571} provides a coreset with a  {\it multiplicative guarantee} that is based on an average sensitivity property of the dataset. 
They provide a lower bound for the size of a multiplicative error coreset proving in particular that in general, it is not possible to achieve $m \ll n$. 
The coreset they build is of cardinality $m \approx \mu\sqrt{nd^3}/\eps^2$ where $\mu \geq 1$ is a complexity measure of the dataset.  
\cite{tolochinsky2018coresets} give a generic multiplicative coreset construction for any monotonic function with $\ell_2^2$ regularization. 
The dependence they get is $\tilde O(d/\eps^2)$ ignoring logarithmic factors. 
Additive approximation coresets are also studies in the $\eps$-approximation literature which is also related to the discrepancy of the problem \cite{DBLP:journals/corr/MustafaV17}. In~\cite{braverman2016new}, some connections are drawn between additive and multiplicative guarantees by providing a method to use an additive guarantee along with sensitivity scores in order to provide a multiplicative guarantee\footnote{These methods might be combined with our results to obtain improved multiplicative guarantees, but this would not be a trivial result and we defer it to future research.}.

@@

Theorems~\ref{thm:streaming11} and~\ref{thm:streaming21} are achieved by deterministic algorithms. 
They could be thought of extensions of the MRL algorithm \cite{MRL} for streaming quantile sketching. 
Quantile sketching falls into this framework since it corresponds to $f(x,q) = 1$ if $x>q$ and $0$ else. The techniques of the above Theorems could also be associated with~\cite{matousek1995approximations}, providing a merge-reduce framework for additive coresets. 
More details and the proof of correctness are given in Appendix ~\ref{app:sketch proof}.

Recently, \cite{DBLP:conf/focs/KarninLL16} provided an improved (optimal) streaming quantile coreset algorithm by improving the merge-reduce technique in a way tailored to the quantile problem. In the typical merge-reduce framework, the algorithm is based on finding an $\eps$-coreset on subsets of size dependent on $\eps$ rather than on $n$. The novelty of \cite{DBLP:conf/focs/KarninLL16} is in suggesting a way to use different values of $\eps$ for these local coreset constructions. This ends up providing a randomized algorithm with no dependence on $n$ and doubly logarithmic dependence on the failure probability. 
%
%In high level, their novelty is using 
Generalizing their construction requires more work and the main ideas are as follows. 
We argued above that $\tilde F_{+}$ and $\tilde F_{-}$ are both good approximations for $F$. 
We can also take $\tilde F_{\pm}$ which is $\tilde F_{+}$ or $\tilde F_{-}$ with equally probability. 
Clearly $|\tilde F_{\pm} - F| \le |E|$ as before. But now, $\E[\tilde F_{\pm}] = F$ as well. 
In the streaming algorithm, we apply this compaction (converting $F$ to $\tilde F_{\pm}$) many times to small subsets of items from the stream. This allows us to use concentration results to bound the overall error. So far, analogous ideas where used in \cite{DBLP:conf/focs/KarninLL16};
the main departure is that $\tilde F_{\pm}$ has half the support of $F$ only in expectation.
